---
title: "Midterm_nba_report"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---
```{r, set-chunk-opts, echo = FALSE, eval = TRUE}
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE
)
```

# Introduction
This report examines per‑36‑minute NBA player statistics (as of March 7, 2025). It covers:

1. Data cleaning and preparation
2. Scoring analysis by position
3. Defensive and rebounding metrics
4. Age and experience relationships
5. Team potential ranking

Scripts are located in `Code/`, and all outputs are written to `Output/`.


# Data Analysis

## Coder1:


## Coder2:



## Coder3:



## Coder4:


## Coder 5: Identifying top 5 potential NBA teams
Based on the analysis for individual players, we can now conduct data analysis at the team level. Upon reviewing the original dataset, we observe that our NBA dataset includes player-level statistics, such as:

1. Youth (Age): Represented by the "Age" column.

2. Scoring Efficiency: Represented by FG%, 3P%, FT%, or PTS.

3. Defensive Ability: Estimated using STL, BLK, and DRB.

By assigning weights to each of these categories—Youth (40%), Scoring Efficiency (30%), and Defensive Ability (30%)—we can calculate the Potential_Score for each team. This allows us to determine the top 6 teams with the highest potential.

Here are the results:

```{r, table-top6_teams_table}
table_one <- readRDS(
  file = here::here("Output/top6_teams_table.rds")
)
table_one
```

We also generate radar plots for these 6 teams using a 9-dimension profile: Youth (inverse age), FG% (Field Goal %), 3P% (Three Point %), FT% (Free Throw %), PTS (Points per game), STL (Steals), BLK (Blocks) and DRB (Defensive rebounds).

```{r, radar-plot}
#| fig.align = "center",
knitr::include_graphics(
  here::here("Output/top6_team_radar_grid.png")
)
```
#Discussion



